Factors affecting reliablity of state-space age-structured assessment models

### 2024-01-26 Project_0/paper/paper.Rmd
edits to project0 paper

We conducted a simulation study where we simulated process errors and observations for 72 operating models with alternative assumptions about fishing history, degree of uncertainty in index and age composition observations, type (recruitment, survival, fishery selectivity, catchability and natural mortality), degree of variation, and correlation of process errors. We fit 20 different estimating models to each of 100 simulated set of observations with alternative assumptions type and correlation structure of process errors, (mean) natural mortality was known or estimated, and a B-H stock recruit relationship was assumed or not.

@@

Bias in spawning stock biomass estimation was generally low for estimating models that assumed the correct source of process error when there was lower observation error. Reliable estimation of stock-recruit relationship parameters only appears possible in ideal situations with lower observation errors in age composition and indices, lower variability in recruitment process errors and large contrast in spawning biomass over time. We found little evidence of bias for many OM process error assumptions when there was contrast in fishing pressure even when there was greater observation error although it can lead to more variable estimates of natural mortality. For OMs where there was bias in natural mortality due to high observation error, estimating the stock-recruit relationship seemed to remove the bias. However, estimation of natural mortality can cause large differences between the true and estimated SSB (that may be unbiased on average) when there is less contrast in fishing pressure over time and higher observation error.

Retrospective patterns were generally weak for all estimation models regardless of the true source of process error, but they can be expected for recruitment even for the correct process error assumptions when observation error is high. When models did exhibit some retrospective pattern, estimating the mean natural mortality rate tended to remove it.

@@

Here we conduct a simulation study with operating models varying by degree of observation error uncertainty, source and variability of process error, and fishing history. The simulations from these operating models are fitted with estimating models that make alternative assumptions for sources of process error, whether a stock-recruit model was estimated, and whether M is estimated. We evaluate whether AIC can correctly determine the correct source of process error and stock effects on recruitment. We also evaluate the degree of bias in the outputs of the assessment model that are important for management.

@@

We used the Woods Hole Assessment Model (WHAM) to configure operating and estimation models in our simulation study \citep[]{millerstock20,stockmiller21}. WHAM is an R package freely available as a github repository. For this study we used [version 1.0.6.9000, commit 77bbd94](https://github.com/timjmiller/wham/tree/77bbd946e4881216a439933473d1c58b21c270c3). This packages has also been used to configure operating and estimating models for closed loop simulations evaluating index-based assessment methods \citep{legaultetal23} and is used for management of haddock, butterfish, American plaice, bluefish, Atlantic cod, and black sea bass in the Northeast US.

We completed a simulation study with a number of operating models that can be categorized based on where process error random effects are assumed: abundance at age (R, R+S), natural mortality (R+M), fleet selectivity (R+Sel), or index catchability (R+q). For each operating model assumptions about variance of process errors and observations are required and the values we used were based on a review of the range of estimates from recent applications of WHAM in management of stocks of haddock, butterfish, and American plaice in the NE US.

In total, we configured 72 operating models with alternative assumptions about the source and variability of process errors, level of observation error in indices and age composition data, and contrast in fishing pressure over time. We fitted 20 estimating models to observations from each of 100 simulations where process errors were also simulated. Estimating models made alternative assumptions about the source of process errors and whether natural mortality (or the mean for models with process error in natural mortality) was estimated and whether a Beverton-Holt stock recruit relationship was estimated within the estimation model. Details of each of the operating and estimating models are described below.

We did not use the bias-correction feature for process errors or observations described by \citet{stockmiller21} for operating and estimating models.Simulations and model fitting were all carried out on the University of Massachusetts Green High-Performance Computing Cluster. All code we used to perform the simulation study and summarize results can be found at https://github.com/timjmiller/SSRTWG/tree/main/Project_0/code.

@@

The population model tracks 10 age classes: ages 1 to 10+ and we assume spawning occurs 1/4 of the way through the year. The maturity at age was a logistic curve with $a_{50}$ = 2.89 and slope = 0.88 and assumed known in all estimation models (Figure \ref{om_maturity}).

@@

We assumed a Beverton-Holt stock recruit function with constant pre-recruit mortality parameters for all operating models. All post-recruit productivity components are constant in the NAA and survey catchability process error operating models. Therefore steepness and unfished recruitment are also constant over the time period for those operating models \citep{millerbrooks21}. We specified unfished recruitment = $R_0 = e^{10}$ and $\Fmsy = F_{40\%} = 0.348$ equated to a steepness of 0.69 and $\alpha=0.60$ and $\beta = 2.4 \times 10^{-5}$ for the
\[
N_{1,y} = \frac{\alpha \text{SSB}_{y-1}}{1 + \beta \text{SSB}_{y-1}} 
\]
Beverton-Holt parameterization (Figure \ref{om_sr}). For OMs without process errors on natural mortality we assumed the rate was assumed 0.2. For OMs with process errors on natural mortality the mean rate was 0.2.

Two alternative fishing histories were used for operating models. In the first scenario, the stock experiences overfishing for the first 20 years and fishing at \Fmsy for the last 20 years. In the second scenario, the stock is fished at \Fmsy for the entire time period. The magnitude of the overfishing assumptions is based on average estimates of overfishing for NE groundfish stocks from \citet{wiedenmannetal19}. \Citet{legaultetal23} also used similar approaches to defining fishing mortality histories for operating models.

@@

For operating models with time-varying random effects for M, steepness is not constant, but we used the same alpha and beta parameters as other operating models this equates to a steepness and R0 at the mean of the time series process for M. For operating models with time-varying random effects for fishery selectivity, \Fmsy is also not constant however we use the same F history as other operating models which corresponds to Fmsy at the mean selectivity parameters.

@@

Standard deviation for log-aggregate catch was 0.1. There were two levels of observation error variance for indices and age composition for both indices and fleet catch. A low uncertainty specification assumed standard deviation of both series of log-aggregate index observations was 0.1 and the standard deviation of the logistic-normal for age composition observations was 0.3 In the high uncertainty specification the standard deviation for log-aggregate indices was 0.4 and that for the age composition observations was 1.5. For all estimating models, standard deviation for log-aggregate observations was assumed known whereas that for the logistic-normal age composition observations was estimated.

@@

The factorial combinations of fishing histories, observation error assumptions, and marginal variance assumptions for recruitment and survival results in 24 different operating models with random effects on recruitment and survival (Table \ref{naa_om_table}).

@@

The factorial combinations of fishing histories, observation error assumptions, and marginal variance and autocorrelation assumptions for natural mortality results in 16 different operating models with random effects on natural mortality (Table \ref{M_om_table}). Internally, WHAM treats natural mortality as a log-transformed parameter 
$$
\log M_y = \mu_M + \epsilon_{M,y}
$$
that is a linear combination of a mean $\mu_M$ and any annual random effects marginally distributed as $\epsilon_{M,y} \sim \text{N}\left(0,\sigma_M^2\right)$. The values for the marginal variance and correlation parameters assumed in the operating models are given in Table \ref{M_om_table}.

@@

The factorial combinations of fishing histories, observation error assumptions, and marginal variance and autocorrelation assumptions for selectivity parameters results in 16 different operating models with random effects on selectivity (Table \ref{sel_om_table}). Internally, WHAM treats selectivity parameter $s$ as a logit-transformed parameter 
$$
\log\left(\frac{p_{s,y}-l_{s}}{u_{s}-p_{s,y}}\right) = \mu_s + \epsilon_{s,y}
$$
that is a linear combination of a mean $\mu_s$ and any annual random effects marginally distributed as $\epsilon_{s,y} \sim \text{N}\left(0,\sigma_s^2\right)$ where the lower and upper bounds of the parameter ($l_s$ and $u_s$) can be specified by the user. All selectivity parameters are either $a_50$ or slope parameters and we assume bounds of 0 and 10 for all selectivity parameters for all operating and estimating models. The values for the marginal variance and correlation parameters assumed in the operating models are given in Table \ref{sel_om_table}.

@@

The factorial combinations of fishing histories, observation error assumptions, and marginal variance and autocorrelation assumptions for catchability results in 16 different operating models with random effects on catchability (Table \ref{q_om_table}). Internally, WHAM treats catchability for index $i$ as a logit-transformed parameter 
$$
\log\left(\frac{q_{i,y}-l_{i}}{u_{i}-q_{i,y}}\right) = \mu_i + \epsilon_{i,y}
$$
that is a linear combination of a mean $\mu_i$ and any annual random effects marginally distributed as $\epsilon_{i,y} \sim \text{N}\left(0,\sigma_i^2\right)$ where the lower and upper bounds of the catchability ($l_i$ and $u_i$) can be specified by the user. Here we assume bounds of 0 and 1000 for all operating and estimating models. For operating and estimation models with process errors on catchability, the temporal variation is only assumed for the first index. The values for the marginal variance and correlation parameters assumed in the operating models are given in Table \ref{q_om_table}

## Estimating models {-}

For each data set simulated from an operating model 20 estimating models were fit. A total of 32 different estimating models were fit across all operating models where the subset of 20 depended on the source of process error (Table \ref{em_table}). The first 20 estimating models in Table \ref{em_table} were fit to simulate data sets from R and R+S operating models. Estimating models 5 to 24 in Table \ref{em_table} were fit to simulate data sets from R+M operating models. Estimating models 5 to 20 and 25-28 in Table \ref{em_table} were fit to simulate data sets from R+Sel operating models. Finally, estimating models 5 to 20 and 29-32 in Table \ref{em_table} were fit to simulate data sets from R+q operating models. The observation error variance of aggregate catch and indices were all assumed known at the true values.

@@

The first measure of reliability we investigated was frequency of convergence when fitting each estimating model to the simulated data sets.
There are various ways to assess convergence of the fit. We summarized 5 alternative categories of convergence.

@@

We measured the frequency of correct model selection using marginal AIC. For a given operating model we the set of models that were considered all made the same assumptions on whether or not to estimate M and whether a stock-recruit function is assumed.

We also measured the frequency of correctly selecting models with Beverton-Holt stock recruit function assumed over models without the stock-recruit function (null model). Because the proportions of simulations where the correct model was chosen for a given set of operating model factors implied generally poor performance of AIC, we analyzed the effect of variation in SSB on the probability of lower AIC of the BH models. We fit binomial generalize linear models to the observed AIC choice for each data set as a function of log(SD(SSB)) and operating model.

@@

We estimated relative bias 
$$
\widehat {\text{RB}}\left(\theta\right) = \frac{1}{N} \sum^N_{i = 1} \frac{\widehat \theta_i - \theta_i}{\theta_i}
$$
of annual SSB and fully-selected fishing mortality for each estimation model and constructed 95\% confidence intervals for the median relative bias using the binomial distribution approach as in \citet{millerhyun18} and \citet{stockmiller21}. We also estimated bias of stock-recruit parameters for estimation models that assumed the Beverton-Holt stock recruit function.

@@

We estimated Mohn's $\rho$ for SSB, fully-selected fishing mortality, and recruitment for each estimation model. We estimated 7 peels for each estimation model. We calculated median 95\% confidence intervals for Mohn's $\rho$ using the same methods as that for relative bias.

@@

R+S EMs fit to R OMs exhibited poor convergence for types 3, 4, and 5 regardless of whether M or a stock-recruit relationship was estimated (Figures \ref{naa_om_em_R_MF_convergence} to \ref{naa_om_em_BH_ME_convergence}). R EMs exhibited high convergence rates for all convergence types for both R and R+S OMs. Convergence rates were high for all convergence types for all EMs in R+S OMs with lower uncertainty in observations. Convergence for most convergence types generally declines most estimating models when mean-log natural mortality rate is estimated and/or a Beverton-Holt stock recruit relationship is estimated even when the process error assumptions of the estimating and operating models match. There can be relatively high convergence probability of type 4 (an invertible hessian) for EMs that do not have the correct process error assumed, but type 5 convergence (no very large standard errors estimated from hessian) for these estimating models is typically much lower.

@@

Marginal AIC accurately determines the correct process error assumptions in estimating models when data are generated from R and R+S OMs, regardless of whether mean log-natural mortality or a stock recruit relationship is estimated (Figures \ref{naa_om_proportion_best_aic_R_MF} to \ref{naa_om_proportion_best_aic_SR_ME}). Adding estimation of mean log-natural mortality or a stock recruit relationship separately has a negligible effect on the accuracy of determining the correct process error assumption. When both are estimated, there is a noticeable reduction in accuracy when OMs have a constant fishing history, observation error is low and largest variability in recruitment process errors.

@@

Marginal AIC most accurately determined the correct source of process error and correlation structure for R+Sel OMs with low observation error (Figures \ref{Sel_om_proportion_best_aic_R_MF} to \ref{Sel_om_proportion_best_aic_SR_ME}). When there was low variability in selectivity process errors and high observation error, R+q or R+S EMs were more likely to have the best AIC. Whether mean log-natural morality or stock recruit relationships were estimated appeared to have little effect on the performance of marginal AIC.
### R+q operating models {-}

@@

## AIC performance for stock-recruit relationship {-}

@@

AIC accuracy for including the B-H stock-recruit relationship was estimated to be highest for models OMs and EMs with process errors only on recruitment and lowest variance of recruitment process errors (Figures \ref{naa_om_MF_BH_glm_AIC_plots} and \ref{naa_om_ME_BH_glm_AIC_plots}). Holding variability in SSB constant, accuracy of stock-recruit  model selection declined with greater variance of recruitment and survival process errors. Simultaneously estimating mean log-natural mortality has a small negative effect on AIC accuracy.

For R+M OMs and EMs there appeared to little effect of correlation in the natural mortality process errors or estimating the mean log-natural mortality on accuracy of AIC in determining a stock-recruit function, holding variability in SSB constant (Figures \ref{M_om_MF_BH_glm_AIC_plots} and \ref{M_om_ME_BH_glm_AIC_plots}).

The range of variation in SSB across simulated populations was lower for R+Sel operating models than R+S or R+M OMs. OMs with contrast in fishing pressure over time is the primary mechanism creating variation in SSB. However, there appears to be little effect of variability or correlation of selectivity process errors or whether mean log-natural mortality was estimated on the accuracy of AIC in selecting the stock recruit relationship (Figures \ref{Sel_om_MF_BH_glm_AIC_plots} and \ref{Sel_om_ME_BH_glm_AIC_plots})

Like the R+Sel OMs, the range of variation in SSB across is only increased by allowing contrast in fishing pressure over time. There appears to be a slight decrease in accuracy in AIC when mean log-natural mortality is estimated and there is a more noticeable increase in accuracy for OMs with lower observation error in those cases. (Figures \ref{q_om_MF_BH_glm_AIC_plots} and \ref{q_om_ME_BH_glm_AIC_plots})

@@

When mean log-natural mortality is known, all estimating models regardless of process error assumptions produce unbiased estimates of SSB over the entire time period for R OMs (Figure \ref{naa_om_em_R_MF_relbias_ssb}). For R+S OMs, only R+S EMs produce unbiased estimation of SSB and bias resulting from other EMs increased with larger variability in survival random effects. Estimating mean log-natural mortality generally produced greater variability in relative errors of SSB estimates and a change in sign of bias for EMs assuming process errors other than R and R+S (Figure \ref{naa_om_em_R_ME_relbias_ssb}). Estimating a stock-recruit relationship had little discernable effect on bias of SSB estimation (Figures \ref{naa_om_em_BH_MF_relbias_ssb}), but in combination with estimating mean log-natural mortality resulted in large bias of R+S EMs for R OMs with high observation error, constant fishing pressure and large variability in recruitment process errors \ref{naa_om_em_BH_ME_relbias_ssb}).

@@

When mean log-natural mortality is known, SSB estimation bias was small or non-existent for all EM process error assumptions when the OMs had low variability of the natural mortality process errors except a tendency toward negative bias in the terminal years for OMs with high observation error and a change in fishing pressure over time (Figure \ref{M_om_em_R_MF_relbias_ssb}). Errors in SSB estimation were most variable with high variability and autocorrelation of natural mortality process errors, but confidence intervals generally covered 0 indicating no evidence of bias. Estimating mean log-natural mortality results in large variability in SSB relative errors, particularly for OMs with high observation error and constant fishing pressure and all EMs produced biased estimation of SSB over all years when OMs had high observation error, a change in fishing pressure and low variability of natural mortality process errors (Figure \ref{M_om_em_R_ME_relbias_ssb}). Like R and R+S OMs, estimating a stock-recruit relationship had little discernible effect on SSB bias (Figures \ref{M_om_em_BH_MF_relbias_ssb} and \ref{M_om_em_BH_ME_relbias_ssb})

@@

Among the process error configurations we used in operating models, we found AIC to be accurate for process errors on recruitment and survival. Fitting models to other OMs rarely preferred R+S EMs and R and R+S EMs were nearly always selected for the matching OMs. For other sources of process error, accuracy of AIC was improved to useful levels when there was larger variability in the process errors and/or lower observation error. Therefore, when data are relatively good and an there is large variability in the unknown process errors, AIC is likely to perform well.

@@

Higher observation error with variation in fishing pressure can produce retrospective patterns even when the the process error assumption is correct, but this retrospective pattern can be reduced if M or the correct stock recruit relationship is estimated. Retrospective patterns tended to be more variable for recruitment and can be large even when the estimating model is correct. Therefore, we recommend emphasis on inspection of retrospective patterns primarily for SSB and F, but further research on retrospective patterns in other assessment model parameters and derived output may be beneficial.

>>>>

We conducted a simulation study where we simulated process errors and observations for 72 operating models with alternative assumptions about fishing history, degree of uncertainty in index and age composition observations, type (recruitment, survival, fishery selectivity, catchability and natural mortality), degree of variation, and correlation of process errors. We fit 20 different estimation models to each of 100 simulated set of observations with alternative assumptions type and correlation structure of process errors, (mean) natural mortality was known or estimated, and a B-H stock recruit relationship was assumed or not.

@@

Bias in spawning stock biomass estimation was generally low for estimation models that assumed the correct source of process error when there was lower observation error. Reliable estimation of stock-recruit relationship parameters only appears possible in ideal situations with lower observation errors in age composition and indices, lower variability in recruitment process errors and large contrast in spawning biomass over time. We found little evidence of bias for many OM process error assumptions when there was contrast in fishing pressure even when there was greater observation error although it can lead to more variable estimates of natural mortality. For OMs where there was bias in natural mortality due to high observation error, estimation the stock-recruit relationship seemed to remove the bias. However, estimation of natural mortality can cause large differences between the true and estimated SSB (that may be unbiased on average) when there is less contrast in fishing pressure over time and higher observation error.

Retrospective patterns were generally weak for all estimation models regardless of the true source of process error, but they can be expected for recruitment even for the correct process error assumptions when observation error is high. When models did exhibit some retrospective pattern, estimation the mean natural mortality rate tended to remove it.

@@

Here we conduct a simulation study with operating models (OMs) varying by degree of observation error uncertainty, source and variability of process error, and fishing history. The simulations from these OMs are fitted with estimation models (EMs) that make alternative assumptions for sources of process error, whether a stock-recruit model was estimated, and whether M is estimated. We evaluate whether AIC can correctly determine the correct source of process error and stock effects on recruitment. We also evaluate the degree of bias in the outputs of the assessment model that are important for management.

@@

We used the Woods Hole Assessment Model (WHAM) to configure operating and EMs in our simulation study \citep[]{millerstock20,stockmiller21}. WHAM is an R package freely available as a github repository. For this study we used [version 1.0.6.9000, commit 77bbd94](https://github.com/timjmiller/wham/tree/77bbd946e4881216a439933473d1c58b21c270c3). This packages has also been used to configure operating and EMs for closed loop simulations evaluating index-based assessment methods \citep{legaultetal23} and is used for management of haddock, butterfish, American plaice, bluefish, Atlantic cod, and black sea bass in the Northeast US.

We completed a simulation study with a number of OMs that can be categorized based on where process error random effects are assumed: abundance at age (R, R+S), natural mortality (R+M), fleet selectivity (R+Sel), or index catchability (R+q). For each OM assumptions about variance of process errors and observations are required and the values we used were based on a review of the range of estimates from recent applications of WHAM in management of stocks of haddock, butterfish, and American plaice in the NE US.

In total, we configured 72 OMs with alternative assumptions about the source and variability of process errors, level of observation error in indices and age composition data, and contrast in fishing pressure over time. We fitted 20 EMs to observations from each of 100 simulations where process errors were also simulated. EMs made alternative assumptions about the source of process errors and whether natural mortality (or the mean for models with process error in natural mortality) was estimated and whether a Beverton-Holt stock recruit relationship was estimated within the EM. Details of each of the operating and EMs are described below.

We did not use the bias-correction feature for process errors or observations described by \citet{stockmiller21} for operating and EMs. Simulations and model fitting were all carried out on the University of Massachusetts Green High-Performance Computing Cluster. All code we used to perform the simulation study and summarize results can be found at https://github.com/timjmiller/SSRTWG/tree/main/Project_0/code.

@@

The population model tracks 10 age classes: ages 1 to 10+ and we assume spawning occurs 1/4 of the way through the year. The maturity at age was a logistic curve with $a_{50}$ = 2.89 and slope = 0.88 and assumed known in all EMs (Figure \ref{om_maturity}).

@@

We assumed a Beverton-Holt stock recruit function with constant pre-recruit mortality parameters for all OMs. All post-recruit productivity components are constant in the NAA and survey catchability process error OMs. Therefore steepness and unfished recruitment are also constant over the time period for those OMs \citep{millerbrooks21}. We specified unfished recruitment = $R_0 = e^{10}$ and $\Fmsy = F_{40\%} = 0.348$ equated to a steepness of 0.69 and $\alpha=0.60$ and $\beta = 2.4 \times 10^{-5}$ for the
\[
N_{1,y} = \frac{\alpha \text{SSB}_{y-1}}{1 + \beta \text{SSB}_{y-1}} 
\]
Beverton-Holt parameterization (Figure \ref{om_sr}). For OMs without process errors on natural mortality we assumed the rate was assumed 0.2. For OMs with process errors on natural mortality the mean rate was 0.2.

Two alternative fishing histories were used for OMs. In the first scenario, the stock experiences overfishing for the first 20 years and fishing at \Fmsy for the last 20 years. In the second scenario, the stock is fished at \Fmsy for the entire time period. The magnitude of the overfishing assumptions is based on average estimates of overfishing for NE groundfish stocks from \citet{wiedenmannetal19}. \Citet{legaultetal23} also used similar approaches to defining fishing mortality histories for OMs.

@@

For OMs with time-varying random effects for M, steepness is not constant, but we used the same alpha and beta parameters as other OMs this equates to a steepness and R0 at the mean of the time series process for M. For OMs with time-varying random effects for fishery selectivity, \Fmsy is also not constant however we use the same F history as other OMs which corresponds to Fmsy at the mean selectivity parameters.

@@

Standard deviation for log-aggregate catch was 0.1. There were two levels of observation error variance for indices and age composition for both indices and fleet catch. A low uncertainty specification assumed standard deviation of both series of log-aggregate index observations was 0.1 and the standard deviation of the logistic-normal for age composition observations was 0.3 In the high uncertainty specification the standard deviation for log-aggregate indices was 0.4 and that for the age composition observations was 1.5. For all EMs, standard deviation for log-aggregate observations was assumed known whereas that for the logistic-normal age composition observations was estimated.

@@

The factorial combinations of fishing histories, observation error assumptions, and marginal variance assumptions for recruitment and survival results in 24 different OMs with random effects on recruitment and survival (Table \ref{naa_om_table}).

@@

The factorial combinations of fishing histories, observation error assumptions, and marginal variance and autocorrelation assumptions for natural mortality results in 16 different OMs with random effects on natural mortality (Table \ref{M_om_table}). Internally, WHAM treats natural mortality as a log-transformed parameter 
$$
\log M_y = \mu_M + \epsilon_{M,y}
$$
that is a linear combination of a mean $\mu_M$ and any annual random effects marginally distributed as $\epsilon_{M,y} \sim \text{N}\left(0,\sigma_M^2\right)$. The values for the marginal variance and correlation parameters assumed in the OMs are given in Table \ref{M_om_table}.

@@

The factorial combinations of fishing histories, observation error assumptions, and marginal variance and autocorrelation assumptions for selectivity parameters results in 16 different OMs with random effects on selectivity (Table \ref{sel_om_table}). Internally, WHAM treats selectivity parameter $s$ as a logit-transformed parameter 
$$
\log\left(\frac{p_{s,y}-l_{s}}{u_{s}-p_{s,y}}\right) = \mu_s + \epsilon_{s,y}
$$
that is a linear combination of a mean $\mu_s$ and any annual random effects marginally distributed as $\epsilon_{s,y} \sim \text{N}\left(0,\sigma_s^2\right)$ where the lower and upper bounds of the parameter ($l_s$ and $u_s$) can be specified by the user. All selectivity parameters are either $a_50$ or slope parameters and we assume bounds of 0 and 10 for all selectivity parameters for all operating and EMs. The values for the marginal variance and correlation parameters assumed in the OMs are given in Table \ref{sel_om_table}.

@@

The factorial combinations of fishing histories, observation error assumptions, and marginal variance and autocorrelation assumptions for catchability results in 16 different OMs with random effects on catchability (Table \ref{q_om_table}). Internally, WHAM treats catchability for index $i$ as a logit-transformed parameter 
$$
\log\left(\frac{q_{i,y}-l_{i}}{u_{i}-q_{i,y}}\right) = \mu_i + \epsilon_{i,y}
$$
that is a linear combination of a mean $\mu_i$ and any annual random effects marginally distributed as $\epsilon_{i,y} \sim \text{N}\left(0,\sigma_i^2\right)$ where the lower and upper bounds of the catchability ($l_i$ and $u_i$) can be specified by the user. Here we assume bounds of 0 and 1000 for all operating and EMs. For operating and EMs with process errors on catchability, the temporal variation is only assumed for the first index. The values for the marginal variance and correlation parameters assumed in the OMs are given in Table \ref{q_om_table}

## Estimation models {-}

For each data set simulated from an OM 20 EMs were fit. A total of 32 different EMs were fit across all OMs where the subset of 20 depended on the source of process error (Table \ref{em_table}). The first 20 EMs in Table \ref{em_table} were fit to simulate data sets from R and R+S OMs. EMs 5 to 24 in Table \ref{em_table} were fit to simulate data sets from R+M OMs. EMs 5 to 20 and 25-28 in Table \ref{em_table} were fit to simulate data sets from R+Sel OMs. Finally, EMs 5 to 20 and 29-32 in Table \ref{em_table} were fit to simulate data sets from R+q OMs. The observation error variance of aggregate catch and indices were all assumed known at the true values.

@@

The first measure of reliability we investigated was frequency of convergence when fitting each EM to the simulated data sets.
There are various ways to assess convergence of the fit. We summarized 5 alternative categories of convergence.

@@

We measured the frequency of correct model selection using marginal AIC. For a given OM we the set of models that were considered all made the same assumptions on whether or not to estimate M and whether a stock-recruit function is assumed.

We also measured the frequency of correctly selecting models with Beverton-Holt stock recruit function assumed over models without the stock-recruit function (null model). Because the proportions of simulations where the correct model was chosen for a given set of OM factors implied generally poor performance of AIC, we analyzed the effect of variation in SSB on the probability of lower AIC of the BH models. We fit binomial generalize linear models to the observed AIC choice for each data set as a function of log(SD(SSB)) and OM.

@@

We estimated bias as the median of the relative errors across all simulations for a given OM and EM combination.
$$
\text{RE}\left(\theta_i\right) = \frac{\widehat \theta_i - \theta_i}{\theta_i}
$$
of annual SSB and fully-selected fishing mortality for each EM and constructed 95\% confidence intervals for the median relative bias using the binomial distribution approach as in \citet{millerhyun18} and \citet{stockmiller21}. We also estimated bias of stock-recruit parameters for EMs that assumed the Beverton-Holt stock recruit function and the mean natural mortality rate parameter when estimated.

@@

We estimated Mohn's $\rho$ for SSB, fully-selected fishing mortality, and recruitment for each EM. We estimated 7 peels for each EM. We calculated median 95\% confidence intervals for Mohn's $\rho$ using the same methods as that for relative bias.

@@

R+S EMs fit to R OMs exhibited poor convergence for types 3, 4, and 5 regardless of whether M or a stock-recruit relationship was estimated (Figures \ref{naa_om_em_R_MF_convergence} to \ref{naa_om_em_BH_ME_convergence}). R EMs exhibited high convergence rates for all convergence types for both R and R+S OMs. Convergence rates were high for all convergence types for all EMs in R+S OMs with lower uncertainty in observations. Convergence for most convergence types generally declines most EMs when mean-log natural mortality rate is estimated and/or a Beverton-Holt stock recruit relationship is estimated even when the process error assumptions of the estimation and OMs match. There can be relatively high convergence probability of type 4 (an invertible hessian) for EMs that do not have the correct process error assumed, but type 5 convergence (no very large standard errors estimated from hessian) for these EMs is typically much lower.

@@

Marginal AIC accurately determines the correct process error assumptions in EMs when data are generated from R and R+S OMs, regardless of whether mean log-natural mortality or a stock recruit relationship is estimated (Figures \ref{naa_om_proportion_best_aic_R_MF} to \ref{naa_om_proportion_best_aic_SR_ME}). Adding estimation of mean log-natural mortality or a stock recruit relationship separately has a negligible effect on the accuracy of determining the correct process error assumption. When both are estimated, there is a noticeable reduction in accuracy when OMs have a constant fishing history, observation error is low and largest variability in recruitment process errors.

@@

Marginal AIC most accurately determined the correct source of process error and correlation structure for R+Sel OMs with low observation error (Figures \ref{Sel_om_proportion_best_aic_R_MF} to \ref{Sel_om_proportion_best_aic_SR_ME}). When there was low variability in selectivity process errors and high observation error, R+q or R+S EMs were more likely to have the best AIC. Whether mean log-natural morality or stock recruit relationships were estimated appeared to have little effect on the performance of marginal AIC.

@@

## AIC performance for the stock-recruit relationship {-}

@@

Among R and R+S OMs, AIC accuracy for including the B-H stock-recruit relationship was estimated to be highest for R OMs and EMs and lowest variance of recruitment process errors (Figures \ref{naa_om_MF_BH_glm_AIC_plots} and \ref{naa_om_ME_BH_glm_AIC_plots}). FOr R+S OMs and holding variability in SSB constant, accuracy of stock-recruit  model selection declined with greater variance of recruitment and survival process errors. Simultaneously estimating mean log-natural mortality has a small negative effect on AIC accuracy, but there is an improvement in AIC accuracy with lower observation error in indices and age composition and lower variation of survival process errors.

For R+M OMs and EMs there appeared to be little effect of correlation in the natural mortality process errors or estimating the mean log-natural mortality on accuracy of AIC in determining a stock-recruit function, holding variability in SSB constant (Figures \ref{M_om_MF_BH_glm_AIC_plots} and \ref{M_om_ME_BH_glm_AIC_plots}). However, there was an increase in accuracy with lower observation error for OMs with higher variability in natural mortality process errors whether the mean natural mortality rate parameter was estimated or not.

The range of variation in SSB across simulated populations was lower for R+Sel OMs than R+S or R+M OMs. OMs with contrast in fishing pressure over time is the primary mechanism creating variation in SSB. However, there appears to be little effect of variability or correlation of selectivity process errors or whether mean log-natural mortality was estimated on the accuracy of AIC in selecting the stock recruit relationship (Figures \ref{Sel_om_MF_BH_glm_AIC_plots} and \ref{Sel_om_ME_BH_glm_AIC_plots}). However, there was an increase in AIC accuracy with lower observation error for OMs with independent process errors whether the mean natural mortality rate parameter was estimated or not.

Like the R+Sel OMs, the range of variation in SSB across simulated populations for R+q OMs is only increased by allowing contrast in fishing pressure over time. There appears to be a slight decrease in accuracy in AIC when mean log-natural mortality is estimated and there is a more noticeable increase in accuracy for OMs with lower observation error in those cases. (Figures \ref{q_om_MF_BH_glm_AIC_plots} and \ref{q_om_ME_BH_glm_AIC_plots}). There was also an increase in AIC accuracy with lower observation error for when the mean natural mortality rate parameter was estimated.

@@

When mean log-natural mortality is known, all estimating models regardless of process error assumptions produce unbiased estimates of SSB over the entire time period for R OMs (Figure \ref{naa_om_em_R_MF_relbias_ssb}). For R+S OMs, only R+S EMs produce unbiased estimation of SSB and bias resulting from other EMs increased with larger variability in survival random effects. Estimating mean log-natural mortality generally produced greater variability in relative errors of SSB estimates and a change in sign of bias for EMs assuming process errors other than R and R+S (Figure \ref{naa_om_em_R_ME_relbias_ssb}). Estimating a stock-recruit relationship had little discernable effect on bias of SSB estimation (Figures \ref{naa_om_em_BH_MF_relbias_ssb}), but in combination with estimating mean log-natural mortality resulted in large bias of R+S EMs for R OMs with high observation error, constant fishing pressure and large variability in recruitment process errors (Figure \ref{naa_om_em_BH_ME_relbias_ssb}).

@@

When mean log-natural mortality is known, SSB estimation bias was small or non-existent for all EM process error assumptions when the OMs had low variability of the natural mortality process errors except a tendency toward negative bias in the terminal years for OMs with high observation error and a change in fishing pressure over time (Figure \ref{M_om_em_R_MF_relbias_ssb}). Errors in SSB estimation were most variable with high variability and autocorrelation of natural mortality process errors, but confidence intervals generally covered 0 indicating no evidence of bias. Estimating mean log-natural mortality results in large variability in SSB relative errors, particularly for OMs with high observation error and constant fishing pressure and all EMs produced biased estimation of SSB over all years when OMs had high observation error, a change in fishing pressure and low variability of natural mortality process errors (Figure \ref{M_om_em_R_ME_relbias_ssb}). Like R and R+S OMs, estimating a stock-recruit relationship had little discernible effect on SSB bias (Figures \ref{M_om_em_BH_MF_relbias_ssb} and \ref{M_om_em_BH_ME_relbias_ssb}).

@@

Among the process error configurations we used in OMs, we found AIC to be accurate for process errors on recruitment and survival. Fitting models to other OMs rarely preferred R+S EMs and R and R+S EMs were nearly always selected for the matching OMs. For other sources of process error, accuracy of AIC was improved to useful levels when there was larger variability in the process errors and/or lower observation error. Therefore, when data are relatively good and an there is large variability in the unknown process errors, AIC is likely to perform well.

@@

Higher observation error with variation in fishing pressure can produce retrospective patterns even when the the process error assumption is correct, but this retrospective pattern can be reduced if M or the correct stock recruit relationship is estimated. Retrospective patterns tended to be more variable for recruitment and can be large even when the EM is correct. Therefore, we recommend emphasis on inspection of retrospective patterns primarily for SSB and F, but further research on retrospective patterns in other assessment model parameters and derived output may be beneficial.
